Study of the hadronization of charged pions

### 2017-01-10 1-Answers-Review.tex

\documentclass[12pt]{article}                                            
\usepackage {epsfig}
\textwidth  16.0cm
\textheight 21.5cm
\topmargin -1.0 cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\pagestyle{plain}
\headheight 0.3cm 
\usepackage[dvipsnames]{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{Myred}{named}{Red}
\pagestyle{plain}
\begin{titlepage}
\begin{flushright}
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~{\today}
\end{flushright}
\vskip 3.cm
\centerline{\bf \color{Myred}{M. Wood (chair), S. Manly, and A. Kim} }
\vskip 0.5cm
\centerline{\bf \LARGE \color{Myred}{Analysis Review of the :} }
\vskip 1.cm
\centerline{\bf \Large ``Study of the Hadronization}
\vskip 0.5cm
\centerline{\bf \Large of Charged Pions”''}
\vskip 0.5cm
\centerline {\bf  by R. Dupre et al.}
\vskip .5cm
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{titlepage}
\setcounter{page}{2}

The committee has reviewed the analysis note and has a number of comments that should be 
addressed.  It has been a long time since the data were taken, and it is an interesting study, so 
we encourage the authors to move forward with the improvements in order to bring it to 
publication.

{\it We would like to thank you for the thorough review and hope you will be satisfied with
our answers.} \\

@@

This report is organized in 3 sections: lager issues to resolve, questions from the reviewers, and 
typographical corrections.  There may be some repetition between the 3 sections.  The notation 
is P\# for page \#.

{\it The answers from the authors of the note are all presented in italic script 
to allow a clear distinction.} \\

@@

The first set of variables is the one actually used for the
analysis and the second set is only used to evaluate the systematic errors, so in the 
note, the formulas are given only for the first set.

The motivation for using such a weight cut based on the product of different unrelated
values is actually directly linked to the inbent 
particles being directed in the beam pipe. The $\pi^-$ acceptance drops very 
strongly at low $p_\perp^2$ leading to very high weights, however there is no problem of 
bin migration
or low statistics because the generated yield is very high in this region. We found that 
the best way
to keep these bins while discarding bins where the weight might be slightly smaller but with 
issues concerning bin migration and statistics was to apply this type of cut.

In term of method, the two
step acceptance correction is hopefully going to be clearer. It is the same as the method used 
in the previous L. El Fassi et al. analysis note. The second step of the procedure is
specifically used to take care of the holes in the acceptance (note, they are not really 
holes because problematic bins are generally on the edges). You can see that the weights
of the previously calculated acceptance are used in this step, ensuring that we do not
double correct the data.

The size of the bins is not constant and is indeed increased in region with small event yield.
However, the multi dimensional binning cannot be fully adapted in a general maner. On this
topic we added tables in the note indicating the bin limits that we used.} \\

@@

14.
P26, bottom, if you have regions with very low acceptance, should they be cut out?

{\it In the case of the $\Delta P_\perp^2$, we need to measure an average of the
transverse momentum distribution. Completely cutting a part of the distribution
would indeed solve the acceptance problem but would make the $\langle p_\perp^2 \rangle$
measurement meaning less. We therefore have to live with it and do our best to
reconstruct the whole distribution even if it means having larger error bars.} \\

15.
P26 - "consistent with HERMES results" - could you add HERMES points to the plot with 
your data points to show what you mean as consistent?

{\it This sentence states that similarly to HERMES, we find that both charged pions have similar
transverse momentum. This can be seen in the Figure 4 of the note where we show the result from 
HERMES. We did not want to mislead the reader into any claim about compatibility between 
HERMES and CLAS data here, so we reworded the sentence. We want to point out that such claim 
would be highly model dependent since it is expected that the effect varies with energy. } \\

@@

16.
P27 fig.22-
 It's unclear to which point the error bar is associated. I think the common 
practice is to slightly shift markers horizontally.

{\it Sorry, we overlooked this for these intermediary plots. TO BE CORRECTED} \\

17.
P27, Fig. 22 - the right column of plots is not the same as those in Fig. 15.  Why change the 
variables?

{\it Sorry, it is a mistake on our part, there is no reason. TO BE CORRECTED. And
add figures after radiative correction!} \\

@@

18.
P28 - What is Coulolmb Correction applied to? Momentum of charged particles? Could you 
plot effect of this correction on particles? Is it really big enough to be visible within our 
resolutions ?

{\it The Coulomb correction is applied to incoming and outgoing electrons as well as
produced charged pions. Indeed the effect is invisible in most of the phase space as
it is comparable to the resolution of the detector (points overlap completely). The 
effect is small but visible only for very low $z$ pions. It has been implemented 
because it was suggested it might be big, but we now find that can be neglected 
at our kinematics. ADD a FIGURE???} \\

@@

19.
P30 - 
fig.24 is missing X axis label. You also may consider 2D histogram for better display.

{\it FIGURE TO BE REPLACED COMPLETELY} \\

20.
P31 - I don’t understand the isospin correction.  It does smooth out the data in Fig. 27; 
however, I am missing the physics motivation.  Why does it matter if the quark came from a 
proton or a neutron?  I think you just clarify the explanation.

{\it This correction is here to take care of the u/d quark imbalance in heavier nuclei.
Indeed the production rate of positive pions on neutrons is significantly reduced and that
should not be mistaken with absorbtion in the nuclear medium. We clarified the text in the
note, hopefully it is now clear.} \\

@@

21.
P33, Sec. 2.5.1 - list the negligible effects.

{\it The sentence has been revised to list all effects tested. } \\

@@

22.
P34, Fig. 28 - why is the resolution poorer than in Fig. 12?

{\it Resolution is the same, it looks wider because on this figure, we show the integrated 
result over the 6 sectors.} \\

@@

23.
P34, why do you have a 1\% normalization error in all the multiplicity ratios when you say just 
above that for this particular sample with two particles in the final state the events in the test 
regions dropped to 0.01?  Doesn’t that mean the target contamination normalization error is 
smaller than 1\%?

{\it Multiplicity ratios are a double ratio. The claim is that the semi inclusive part 
is not affected, but the inclusive electron ratio is affected. This is included
as a normalisation error. We revisited slightly the paragraph to make it clearer.} \\

### 2017-01-10 chap5.tex
Two more answers

We used a 4-dimensional binning to divide the large phase space of our two particles' final 
state. To evaluate systematic errors associated with this correction, we compared the
method using two different sets of kinematic variables, as presented in 
table~\ref{tab:AcceptBinning}. The first set of variables is the one actually used for the
analysis and the second set is only used to evaluate the systematic errors, so in the 
following, the formulas are given only for the first set. The total number of bins is 
constrained by the amount of generated data and must be maintained reasonably low in order 
to keep a good statistical precision and a small bin migration.

@@

@@

Figure \ref{fig:AcceptPlots}, where only statistical errors are 
represented, shows the new results after applying the acceptance correction.
The acceptance correction effects appear to be in the order of $10$\% for 
multiplicity ratios and $30$\% for the transverse momentum broadening. This is a surprisingly significant correction with regard to a few centimeters separating the two targets. The reason for such a large correction is that the small shift in the $z$-vertex position changes slightly the minimum angle of the detector in the region where most hadron production is concentrated, {\it i.e.} at low \pt close to the beam line. The $\pi^-$s were more affected due to the inbending magnetic field. Indeed, the low \pt $\pi^-$s had very low acceptance which led to large correction factors. The large spread in weights makes the errors on \dpt very large, and the related results difficult to exploit. However, our transverse momentum broadening results of both pions matched each other after this correction, which is consistent 
with HERMES results.

@@

The important excess of neutrons in heavy nuclei leads to a modification of both $\pi$ multiplicities on DIS events. Using Hall~C results 
\cite{Asaturyan:2011mq}, shown in figure \ref{fig:IsoSpin}, we evaluated the correction factors for this effect. Our simple estimation is solely based on 
proton and neutron numbers that gave $\pi^+$s results shown in table~\ref{tab:isospin}. As the effect on $\pi^-$s was found to be consistent with zero, no correction was applied to their multiplicity ratios. 
We attribute a 10\% normalization error to the effect of the isospin correction on our results. It is relatively chosen based on the precision of the Hall~C measurement, and the limited phase space coverage of Hall-C data compared to ours.

@@

It is important to mention that the isospin correction is applied only to rates, which means applied only to $\pi^+$s multiplicity ratios. The transverse momentum broadening could, in principle, be affected, however, the results from \cite{Asaturyan:2011mq} showed no isospin effects in \ptp. Therefore, we are confident that our \dpt results should not be sensitive to any isospin effects.

We finally note that the $A$ dependence of multiplicity ratios of both charged pions became parallel after the isospin correction\footnote{Within 
normalization errors presented in section \ref{sec:TotSys}} (see figure 
\ref{fig:IsoPlot}). Considering previous measurements and existing models, 
this result was expected and demonstrated the importance of this correction.

\begin{figure}[tbp]
\centering
\includegraphics[width=8cm] {chap5-fig/c_RvA.png}
\caption {Multiplicity ratios as a function of $A^{1/3}$ 
with only the isospin correction applied.}
\label{fig:IsoPlot}
\end{figure}

@@

  - particles' re-scattering in different detectors,

@@

To evaluate those errors, we are comparing the kinematics of reconstructed particles with the generated ones. Each reconstructed particle is associated with its generated parent if both particles have a similar momentum and angle coverage. However, the precision of the measured kinematical distribution $\Delta p$, $\Delta \theta$ and $\Delta \phi$ were determined as $\Delta x = {\sum_n |x_{gen} - x_{rec}| \over n}$, where $n$ is the kinematical bin number. Using the same simulation as for the acceptance, we found ${\delta p \over p } = 0.03$, $\delta \theta = 3$ mrad and $\delta \phi = 10$~mrad. Nevertheless of being much larger than the nominal CLAS resolutions published in \cite{Mecking:2003zu}, they are still reasonable for our measurement. Although, we evaluated the systematic errors associated with other variables\footnote{These values were evaluated for a typical kinematics range. However, the results can be much larger at extreme kinematics (large \pt or Q$^2$).}, such as $\delta Q^2 \sim 0.013$ GeV$^2$/c$^2$, $\delta z \sim 0.4 \%$, and $\delta P_\perp^2 \sim 0.004$ GeV$^2$/c$^2$. These errors are overall negligible since they are several times smaller than our usual bin sizes (see figures in a section \ref{prelim}).

We also evaluated the particles' misidentification and the re-scattering effect in different detectors or coils. The misidentification usually leads to a high 
contamination on the kinematical distributions tails, which is manifested most in our case on the tail of the \pt distribution. This effect is due to a low probability to produce high \pt events, therefore the contribution from the accidental background is relatively increased. As a consequence, we used the following cut $p_\perp^2 < 2.5$~GeV$^2$/c$^2$, which basically discarded only a very little amount of data (about one $\pi$ from 30000 pions). The misidentification of electrons was found to be too small in the order of $1\,e^-$ each 1000 events, hence does not contribute significantly to this uncertainty. For pions, the main contamination comes from kaons with momenta above 2~GeV/c ($\sim$ 3\% for $\pi^+$s and $\sim$ 0.5\% for $\pi^-$s). Protons also contaminates a $\pi^+$ sample at high momenta (up to a few \%). 
%TODO Justify the pt cut with a figure?

In conclusion, the uncertainty related to particles' misidentification is taken into account in the point-to-point systematic errors by assuming a 100\% contamination level. Other effects were found negligible.

@@

Because of some reconstruction errors or a re-scattering on detectors' materials, it 
is possible to associate a particle with the wrong target. To estimate this 
effect, we looked, in the experimental data, at the number of reconstructed events upstream and downstream our targets, where nothing should be detected. Thus, we defined two test regions (see figure \ref{fig:targetleak}) to evaluate a targets' contamination. The upstream region, named as a test region 1, was given the size of the solid target's selection cut to evaluate the contamination from the liquid target. However, the downstream region, named as a test region 2, covered the same size of the liquid target's selection cut to estimate the leak from the solid target. The distance between detection and test regions is identical to the 
distance between the two targets' detection regions.

We found that the number of electrons in the test region 1 and 2 represents, 
respectively, 1 and 2\% of the total number of events. However, in the case of semi-inclusive measurements, where we request two particles in the final state, this number drastically dropped, and became in the order of 0.01\%. In conclusion, only the number of electrons is significantly affected by a target's contamination issue, which leads to a normalization error of 1\% in all multiplicity ratios.

>>>>

We used a 4-dimensional binning to divide the large phase space of our two particles' final 
state. To evaluate systematic errors associated with this correction, we compared the
method using two different sets of kinematic variables, as presented in 
table~\ref{tab:AcceptBinning} and \ref{tab:AcceptBins}. The first set of variables is the one actually used for the
analysis and the second set is only used to evaluate the systematic errors, so in the 
following, the formulas are given only for the first set. The total number of bins is 
constrained by the amount of generated data and must be maintained reasonably low in order 
to keep a good statistical precision and a small bin migration.

@@

\begin{table}[htbp]
\center
\begin{tabular} {|l|c|c|c|c|}
\hline
 $\nu$ & \xb & $p_\pi$ & $-t$ \\ \hline
 2.20 & 0.12 & 0.25 &-3.50 \\ 
 3.00 & 0.18 & 0.60 &-2.10 \\ 
 3.40 & 0.22 & 1.00 &-1.75 \\ 
 3.75 & 0.27 & 1.45 &-1.50 \\ 
 4.00 & 0.33 & 1.70 &-1.30 \\ 
 4.25 & 0.60 & 2.15 &-1.10 \\ 
      &      & 3.00 &-0.75 \\ 
      &      & 4.25 &-0.00 \\ \hline
\end{tabular}
\quad
\begin{tabular} {|l|c|c|c|c|}
\hline
 $Q^2$ & $\nu$ & $z$ & $-t$ \\ \hline
 1.00 & 2.20 & 0.00 & -3.50 \\
 1.20 & 3.00 & 0.20 & -2.10 \\
 1.40 & 3.40 & 0.40 & -1.75 \\
 1.65 & 3.75 & 0.45 & -1.50 \\
 2.10 & 4.00 & 0.52 & -1.30 \\
 4.00 & 4.25 & 0.62 & -1.10 \\
      &      & 0.80 & -0.75 \\
      &      & 1.00 & -0.00 \\ \hline
\end{tabular}
\caption{Bin limits used for the acceptance correction described in table \ref{tab:AcceptBinning}.}
  \label{tab:AcceptBins}
\end{table}

@@

Figure \ref{fig:AcceptPlots}, where only statistical errors are 
represented, shows the new results after applying the acceptance correction.
The acceptance correction effects appear to be in the order of $10$\% for 
multiplicity ratios and $30$\% for the transverse momentum broadening. This is a surprisingly significant correction with regard to a few centimeters separating the two targets. The reason for such a large correction is that the small shift in the $z$-vertex position changes slightly the minimum angle of the detector in the region where most hadron production is concentrated, {\it i.e.} at low \pt close to the beam line. The $\pi^-$s were more affected due to the inbending magnetic field. Indeed, the low \pt $\pi^-$s had very low acceptance which led to large correction factors. The large spread in weights makes the errors on \dpt very large, and the related results difficult to exploit. However, our transverse momentum broadening 
results of both pions matched each other after this correction, which is similar 
to the HERMES results (see figure~\ref{fig:her3}).

@@

The important excess of neutrons in heavy nuclei leads to a modification of 
both $\pi$ multiplicities regardless of nuclear effects. Using Hall~C results 
\cite{Asaturyan:2011mq}, shown in figure \ref{fig:IsoSpin}, we evaluated the 
correction factors for the imbalance of u/d quarks in heavy nuclei compared
to deuterium. Our simple estimation is solely based on proton and neutron 
numbers and assumes that this isospin correction does not depend on any
kinematic variable. We found that multiplicity ratios of $\pi^+$ need small
corrections, shown in table~\ref{tab:isospin}. The effect on $\pi^-$s was 
found to be consistent with zero, so no correction was applied to their 
multiplicity ratios. We attribute a normalization error to the finale results
equal to 10\% of the isospin correction. This is evaluated based on the 
precision of the Hall~C measurement.

@@

It is important to mention that the isospin correction is only relevant to 
rate measurements. The transverse momentum broadening could, in principle, 
be affected, however, the results from \cite{Asaturyan:2011mq} showed no 
isospin effects in \ptp. Therefore, we are confident that our \dpt results 
should not be sensitive to any isospin effects.

We finally note that the $A$ dependence of multiplicity ratios of both 
charged pions became parallel after the isospin correction\footnote{Within 
normalization errors presented in section \ref{sec:TotSys}} (see figure 
\ref{fig:IsoPlot}). Considering previous measurements and existing models, 
this result was expected and demonstrates the importance of this correction.

\begin{figure}[tbp]
\centering
\includegraphics[width=8.5cm] {chap5-fig/c_RvA.png}
\caption {Multiplicity ratios as a function of $A^{1/3}$ with the isospin correction applied.}
\label{fig:IsoPlot}
\end{figure}

@@

  - particles' scattering in different detectors,

@@

To evaluate these errors, we are comparing the kinematics of reconstructed 
particles with the generated ones. Each reconstructed particle is associated 
with its generated parent if both particles have a similar momentum and 
angle. The precision of the measured kinematical distribution 
$\Delta p$, $\Delta \theta$ and $\Delta \phi$ were determined as $\Delta x = 
{\sum_n |x_{gen} - x_{rec}| \over n}$. Using the same simulation as for 
the acceptance, we found 
${\delta p \over p } = 0.03$, $\delta \theta = 3$ mrad and $\delta \phi = 
10$~mrad. While being larger than the nominal CLAS resolutions published in 
\cite{Mecking:2003zu}, they are still reasonable for our measurement. We also 
evaluated these errors in term of other variables, such as 
$\delta Q^2 \sim 0.013$ GeV$^2$/c$^2$, $\delta z \sim 0.4 \%$ and $\delta 
P_\perp^2 \sim 0.004$ GeV$^2$/c$^2$ (\footnote{These values are evaluated for 
a typical kinematics range. The results can be larger at extreme kinematics 
(large \pt or Q$^2$).}). These errors are overall negligible since they are 
several times smaller than our usual bin sizes (see figures in a section 
\ref{prelim}).

We also evaluated the particles' misidentification and scattering on materials
from the detectors or coils. For this study, we are looking at reconstructed 
particles that do not have a clear parent (either because they are misidentified
or because they scattered and their kinematic changed significantly). The 
misidentification of electrons was found to be small, of the order of 1 in 
1000 $e^-$ in all the kinematic space, hence this issue does not contribute 
significantly to this uncertainty. We found however for pions
a high proportion of accidental events on the kinematical distributions tails, 
in particular of the \pt distribution. This effect is due to the low 
probability to produce high \pt events, leading to an increased relative 
contribution from the accidental background. As a consequence, we used the 
following cut $p_\perp^2 < 2.5$~GeV$^2$/c$^2$, which discards only a very 
little amount of data (about 1 in 30000 $\pi$). On the contamination side,
most events come from kaons with momenta above 2~GeV/c ($\sim$ 3\% for 
$\pi^+$s and $\sim$ 0.5\% for $\pi^-$s). Protons also contaminate the 
$\pi^+$ sample but only for the highest momenta (up to a few \%).

In conclusion, the uncertainty related to particles' misidentification is 
taken into account with the attribution of a point-to-point systematic 
error equal to 100\% of the contamination level. The accidentals are taken 
care of using a cut on the high \pt distribution. Finally, resolution effects
were found to be negligible.

@@

Because of reconstruction errors or scattering on detectors' material, it 
is possible to associate a particle with the wrong target. To estimate this 
effect, we looked, in the experimental data, at the number of reconstructed 
events upstream and downstream of our targets, where nothing should be 
detected. Thus, we defined two test regions (see figure \ref{fig:targetleak}) 
to evaluate the associated contamination between the targets. The upstream 
region, named test region 1, is of the size of the solid target's selection 
cut to evaluate the contamination from the liquid target. The downstream region, 
named test region 2, covers the same size of the liquid target's selection cut 
to estimate the leak from the solid target. The distance between detection and 
test regions is identical to the distance between the two targets' detection 
regions.

We found that the number of electrons in the test region 1 and 2 represents, 
respectively, 1 and 2\% of the total number of events. However, in the case 
of semi-inclusive measurements, where we request two particles coming from the
same spot in the final state, this number drastically dropped to $\sim$0.01\%. 
Therefore, only the number of electrons is significantly affected by the 
target's contamination issue, which is an issue only for multiplicity ratios.
We associate a normalization error of 1\% to all multiplicity ratios to account
with this issue.
